R Notebook

### 2019-06-06 ICS_Cleaning.Rmd
includes and merges new export

```{r Donor Info}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
#Import information about the samples from my lab notes
    PBMC<-read.csv("ICS_PBMC.csv")
#remove samples that are other helminths since we dropped that group
#Not US, US and Unknown are from the Yerkes donors
    PBMC<-dplyr::filter(PBMC, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" |  Helminth.Status=="US" | Helminth.Status== "Unknown")
    PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("US", "Not US", "Negative", "Schisto+", "Unknown"))
#order the levels of TB status to be from least to most exposure/disease
    PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Naïve","BCG","Healthy Control", "LTBI", "Active TB", "Unknown"))
#take out all the stuff that isn't consistent like the "TM" in experiment name and the % in viability
    PBMC$Viability <- gsub("%/%" , "", PBMC$Viability)
    PBMC$Viability<-gsub("%" , "" , PBMC$Viability)
    PBMC$Exp.Date <- gsub("TM", "", PBMC$Exp.Date)
#Some of the samples had weird suffixes that have to be removed to match the epi data  
    PBMC$Study.ID<-gsub("- plasma", "", PBMC$Study.ID)
    PBMC$Study.ID<-gsub("-1", "", PBMC$Study.ID)
    PBMC$Study.ID<-gsub("-0", "", PBMC$Study.ID)
#Remove samples that are being excluded from the ICS analysis
PBMC<- dplyr::filter(PBMC, !(Excluded=="EX"))%>%
    dplyr::rename(TB = TB.Status, SM = Helminth.Status, StudyID = Study.ID)%>% #match the variable names to ICS variable names
    dplyr::select(StudyID, TB, SM, Viability, Cells.Vial, Excluded) #take only the columns you need

#Import the clinical enrollment data from the cleaned merged epi file
    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya-filledin.csv")
#The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
    QFT$StudyID<-tstrsplit(QFT$StudyID, "_")[[1]]
#trim the data set to only those study IDs in the lab notes and select the columns you want
    QFT2<-dplyr::select(QFT, StudyID, age, sex, QFT, SchistosomaIntensity)%>%
        dplyr::filter(StudyID %in% PBMC$StudyID)

#Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
    DonorInfo<-merge(PBMC, QFT2, by="StudyID", all = TRUE)
#The class of variables gets messed up in the merge so re-class them
    DonorInfo$Cells.Vial<-as.numeric(as.character(DonorInfo$Cells.Vial))
    DonorInfo$QFT<-as.numeric(as.character(DonorInfo$QFT))
#Match the variable names of the ICS data
    DonorInfo<-dplyr::rename(DonorInfo, Donor = StudyID)
    DonorInfo$TB<-gsub("Active TB", "TB", DonorInfo$TB)
    DonorInfo$TB<-gsub("Healthy Control", "HC", DonorInfo$TB)
    DonorInfo$SM<-gsub("Negative", "X", DonorInfo$SM)
    DonorInfo$SM<-gsub("Schisto+", "SM", DonorInfo$SM)
    DonorInfo$SM<-gsub("\\+", "", DonorInfo$SM)

#Manual Edits that may get deleted after Kenya script is finalized
    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")
```

@@

l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv, CD4.csv, 
        CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) clean_flowjo_output(x))
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) fix.names(x))
l<-lapply(l, function(x) class_change(x))
l<-lapply(l, function(x) changeTB(x))

@@

```{r change colnames}
#I only bothered to change the names in flowjo for CD4/8 Cytokine exports so the rest are ugly
newnames<-names(CD8_Cytokine.csv)
#luckily the columns are in the same order so I can just take the names from one of those 
names(CD4_Counts.csv)<-newnames[1:21]
#and give it to all the counts tables
names(CD8_Counts.csv)<-newnames
names(GD_Counts.csv)<-newnames
names(GD_Cytokine.csv)<-newnames
```

@@

@@

```{r create new combo cytokine variables}
#If you want to know the double producers without considering the opposite TH cytokines
library(dplyr)
filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) dplyr::mutate(x, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
                   IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x)
)
for (i in 1:length(filenames)) {
    assign(filenames[i], 
           data.frame(l[i]))}
```

@@

```{r write files to "clean" directory}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
df.list<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv,
              CD4.csv, CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv,
              GD_Cytokine.csv)
dfnames<-list("CD4_Counts", "CD4_Cytokine", "CD4_MFI", "CD4_Positive",
              "CD4", "CD8_Counts", "CD8_Cytokine", "GD_Counts",
              "GD_Cytokine")
newfilenames<-paste(dfnames, "clean", date(), ".csv", sep="_")
#write.csv(df.list[1], file= newfilenames[1])
write.csv(CD4_Counts.csv, file = "CD4_Counts_clean.csv")
write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine_clean.csv")
write.csv(CD4_MFI.csv, file = "CD4_MFI_clean.csv")
write.csv(CD4_Positive.csv, file = "CD4_Positive_clean.csv")
write.csv(CD4.csv, file = "CD4_clean.csv")
write.csv(CD8_Counts.csv, file = "CD8_Counts_clean.csv")
write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine_clean.csv")
write.csv(GD_Counts.csv, file = "GD_Counts_clean.csv")
write.csv(GD_Cytokine.csv, file = "GD_Cytokine_clean.csv")
write.csv(DonorInfo, file = "ICS_Donor_Info_clean.csv")
```

>>>>

@@

l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_new.csv,
        CD4_phenotype_new.csv, CD4_phenotype.csv, CD4.csv, 
        CD8_Counts.csv, CD8_Cytokine.csv, 
        GD_Counts.csv, GD_Cytokine.csv
        )
l<-lapply(l, function(x) clean_flowjo_output(x))
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
l<-lapply(l, function(x) fix.names(x))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) remove_donors_ICS(x))
l<-lapply(l, function(x) class_change(x))
l<-lapply(l, function(x) changeTB(x))

@@

```{r change colnames}
#I only bothered to change the names in flowjo for CD4/8 Cytokine exports so the rest are ugly
newnames<-names(CD8_Cytokine.csv)
#luckily the columns are in the same order so I can just take the names from one of those 
names(CD4_Counts.csv)<-newnames[1:21]
#and give it to all the counts tables
names(CD8_Counts.csv)<-newnames
names(GD_Counts.csv)<-newnames
names(GD_Cytokine.csv)<-newnames

CD4_Cytokine.csv <- dplyr::rename(CD4_Cytokine.csv, TH1.Freq = Th1.Freq)
```

@@

#creating variables for TH1 and TH2 that include co-prouducing cells
CD4_Counts.csv$TH1.Total<-(CD4_Counts.csv$TH1 + CD4_Counts.csv$TH1.2)
CD4_Counts.csv$TH2.Total<-(CD4_Counts.csv$TH2 + CD4_Counts.csv$TH1.2)

#Creating a total counts column
CD4_Counts.csv$Total<-CD4_Counts.csv$x_x_x_x + CD4_Counts.csv$Cyt

@@

```{r update and organize cytokine variables}
#If you want to know the double producers without considering the opposite TH cytokines
library(dplyr)
CD4_Cytokine.csv<-dplyr::mutate(CD4_Cytokine.csv, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
                   IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x,
                   TH1.Total = TH1.Freq + TH1.2.Freq,
                   TH2.Total = TH2.Freq + TH1.2.Freq
                   )
```

```{r update and merge phenotype data}
CD4<-merge(CD4.csv, CD4_new.csv, by=c("Sample","Donor","TB","SM","Stim" ), all=TRUE)
CD4_phenotype<-merge(CD4_phenotype.csv, CD4_phenotype_new.csv, by=c("Sample","Donor","TB","SM","Stim" ), all=TRUE)
```

@@

```{r write files to "clean" directory}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
write.csv(CD4_Counts.csv, file = "ICS_CD4_Counts_clean.csv")
write.csv(CD4_Cytokine.csv, file = "ICS_CD4_Cytokine_clean.csv")
write.csv(CD4_MFI.csv, file = "ICS_CD4_MFI_clean.csv")
write.csv(CD4_phenotype, file = "ICS_CD4_phenotype_clean.csv")
write.csv(CD4, file = "ICS_CD4_clean.csv")
write.csv(CD8_Counts.csv, file = "ICS_CD8_Counts_clean.csv")
write.csv(CD8_Cytokine.csv, file = "ICS_CD8_Cytokine_clean.csv")
write.csv(GD_Counts.csv, file = "ICS_GD_Counts_clean.csv")
write.csv(GD_Cytokine.csv, file = "ICS_GD_Cytokine_clean.csv")
```
